---
title: "Tutorial"
author: "Elisabeth Pigott"
date: "11/16/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Install the package.

```{r, eval=FALSE}
devtools::install_github("elisabethpigott/pigott.package")
```

Load the package.

```{r}
library(pigott.package)
```

## Converting a species community dataframe to a matrix for further analysis

This function will convert a species community dataframe and transform it into a matrix. You can then use this matrix to plug into community analysis functions. You must use a dataframe with only Species, Site_names, and Count or Abundance in three separate columns. For my example data, formal site names are irrelevant, so I use Mo_since_burn_grouped instead of Site_names as my independent variable. The output will be a matrix of spread out species community data.

```{r}

library(tidyverse)
bees <- read_csv("../inst/extdata/bee_fxndata.csv")
print(bees)
beematrix(bees, Species, Mo_since_burn_grouped, Count) 

```

## Analysing species diversity (Hill Numbers) using a community data matrix

This function uses iNEXT to report the effective number of species as Hill numbers of the order q. 